### 2020-04-06 paper1/article.tex
more

\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
\bibliographystyle{plainurl}
\title{The Impact of Class Immutability on NCSS}
\author{Yegor Bugayenko}
  {Huawei Technologies Co., Ltd.}
  {Russian Research Institute (RRI)}{Moscow, Russia}{yegor.bugayenko@huawei.com}
\authorrunning{Yegor Bugayenko}
\ccsdesc[100]{Software and its engineering~Object oriented languages}
\keywords{Object-Oriented Programming, Constructors, Cohesion}
\usepackage[utf8]{inputenc}
\usepackage[inline]{enumitem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{pgfplots}
\usepackage{verbatimbox}
\usepackage{interval}
\usepackage{hyperref}
\usepackage{minted}
  \setminted{fontsize=\footnotesize}
  \setminted{breaklines}
  \usemintedstyle{bw}
\newcommand{\code}[1]{\texttt{#1}}
\newenvironment{nicetable}
  {\setlength{\parindent}{0em}\medskip\small}
  {\medskip}
\begin{document}
\maketitle

\begin{abstract}
...
\end{abstract}

\section{Introduction}

One of the most important factors negatively affecting the quality
of object-oriented software is the length of the code. Larger classes
cause readability and maintability issues. One of the simplest way
to calculate size of a Java class is via the NCSS metric, which
stands for Non-Commenting Source Statements. Larger classes with bigger
amount of statements demonstrate higher values of NCSS.

Immutability is a property of a class in OOP. A class is immutable if
it's impossible to modify its attributes after instantiation. In Java this
technically means that all attributes have \texttt{final} modifier attached
to them.

This research empirically validates the relationship between immutability
and NCSS.

\section{Related Work}

No similar research has been done so far.

\section{Empirical Results}

A list of Java repositories were retrieved from GitHub via their
public API. The first 50 repositories were taken, which satisfied
the selection criteria:
\begin{enumerate*}[label={\arabic*)}]
\item more than 1000 stars,
\item more than 200 Kb of data,
\item not archived,
\item public.
\end{enumerate*}
The list included popular Java open source products, like
Spring Framework, RxJava, Google Guava, and others.

Files with \texttt{.java} extension were taken from all repositories.
There were 20,500 files found.

NCSS metric was calculated for each Java class.

Immutability was calculated for each Java class. A class was considered
immutable if it didn't have any attributes without \texttt{final} modified.

The Figure~\ref{fig:1} demonstrates the results obtained. The x-axis is
the value of NCSS. The y-axis is the share of immutable classes among
all classes with the given NCSS.

\begin{figure}[h]
  \input{../ncss.tex}
  \caption{Distribution of classes by NCSS}
  \label{fig:1}
\end{figure}

The tendency is visually obvious: larger classes are less often immutable.

\section{Conclusion}

It was empirically confirmed that larger Java classes tend to be immutable
less frequently than smaller ones.

\end{document}

### 2020-04-06 calc.py

@@ -7,12 +7,16 @@ def immutable(tree):
     raise Exception('This is not a class')
   fields = tlist[0][1].filter(javalang.tree.FieldDeclaration)
   flist = [v for v in fields]
-  if not flist:
-    raise Exception('This are no attributes')
   immutable = True
-  for path, node in fields:
+  found = 0
+  for path, node in flist:
+    if 'static' in node.modifiers:
+      continue
     if not 'final' in node.modifiers:
       immutable = False
+    found += 1
+  if found == 0:
+    raise Exception('There are no attributes')
   return immutable
 
 def ncss(tree):
@@ -30,9 +34,9 @@ def ncss(tree):
   return metric
 
 with open(sys.argv[1], encoding='utf-8') as f:
-  raw = javalang.parse.parse(f.read())
-  tree = raw.filter(javalang.tree.ClassDeclaration)
   try:
+    raw = javalang.parse.parse(f.read())
+    tree = raw.filter(javalang.tree.ClassDeclaration)
     print(str(ncss(raw)) + ',' + ('yes' if immutable(tree) else 'no'))
   except Exception as e:
     sys.exit(e.args)

### 2020-04-06 draw.rb

@@ -1,0 +1,38 @@
+
+data = []
+File.open(ARGV[0]).read.each_line do |l|
+  ncss, flag = l.strip.split(',', 2)
+  ncss = ncss.to_i
+  next if ncss > 2000
+  data << { ncss: ncss, immutable: flag == 'yes' }
+end
+
+steps = 100
+lo = data.map { |r| r[:ncss] }.min
+hi = data.map { |r| r[:ncss] }.max
+step = (hi - lo).to_f / steps
+
+cols = []
+(0..steps - 1).each do |s|
+  all = data.select { |r| r[:ncss] > s * step && r[:ncss] < s * (step + 1) }
+  total = all.count
+  immutable = all.select { |r| r[:immutable] }.count
+  cols << {
+    total: all.count,
+    share: total.zero? ? 0 : immutable.to_f / total
+  }
+end
+
+puts "% Lo: #{lo}"
+puts "% Hi: #{hi}"
+puts "% Step: #{step} (steps: #{steps})"
+
+puts '\begin{tikzpicture}' +
+'\begin{axis}[axis lines=middle, xlabel={NCSS}, ylabel={$n_i$},' +
+"xmin=#{lo}, xmax=#{hi}, ymin=0, ymax=1," +
+"extra tick style={major grid style=black}," +
+'grid=major]\addplot [only marks, mark size=1pt] table {' + "\n"
+cols.each_with_index do |v, i|
+  puts "  #{lo + i * step} #{v[:share]}\n"
+end
+puts '};\end{axis}\end{tikzpicture}'

### 2020-04-06 test/a.java

@@ -1,0 +1,3 @@
+class Foo {
+  public static int TEST = 1;
+}

### 2020-04-06 test/b.java

@@ -1,0 +1,4 @@
+class Foo {
+  private int x;
+  private final int y;
+}

### 2020-04-06 test/c.java

@@ -1,0 +1,4 @@
+class Foo {
+  private final int x;
+  private final int y;
+}

### 2020-04-06 paper1/article.tex
more

@@
-\documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
+% \documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
+\documentclass[12pt]{article}
 \bibliographystyle{plainurl}
 \title{The Impact of Class Immutability on NCSS}
-\author{Yegor Bugayenko}
-  {Huawei Technologies Co., Ltd.}
-  {Russian Research Institute (RRI)}{Moscow, Russia}{yegor.bugayenko@huawei.com}
-\authorrunning{Yegor Bugayenko}
-\ccsdesc[100]{Software and its engineering~Object oriented languages}
-\keywords{Object-Oriented Programming, Constructors, Cohesion}
+\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\Moscow, Russia\\yegor.bugayenko@huawei.com}{}{}
+% \authorrunning{Yegor Bugayenko}
+% \ccsdesc[100]{Software and its engineering~Object oriented languages}
+% \keywords{Object-Oriented Programming, Immutability, NCSS}
 \usepackage[utf8]{inputenc}
@@
   {\medskip}
+\input{../total}
 \begin{document}
+\raggedbottom
 \maketitle
@@
 A list of Java repositories were retrieved from GitHub via their
-public API. The first 50 repositories were taken, which satisfied
+public API. The first \totalrepos{} repositories were taken, which satisfied
 the selection criteria:
@@
 Files with \texttt{.java} extension were taken from all repositories.
-There were 20,500 files found.
+There were \totaljavafiles{} files found. Classes without a single
+non-\texttt{static} attribute were excluded (including interfaces and enums).
 
-NCSS metric was calculated for each Java class.
-
-Immutability was calculated for each Java class. A class was considered
-immutable if it didn't have any attributes without \texttt{final} modified.
+NCSS metric and immutability were calculated for each Java class.
+A class was considered immutable if it didn't have any
+attributes without \texttt{final} modified.
 
@@
 the value of NCSS. The y-axis is the share of immutable classes among
-all classes with the given NCSS.
+all classes with the given NCSS. The diameter of the plot on the graph
+is related to the amount of classes found for the specific NCSS.
 
@@
   \input{../ncss.tex}
-  \caption{Distribution of classes by NCSS}
+  \caption{Distribution of class immutability by NCSS}
   \label{fig:1}
@@
 
-The tendency is visually obvious: larger classes are less often immutable.
+The tendency is visually obvious: larger classes are less often immutable. It
+is also obvious that the majority of classes have NCSS values smaller than 100.
 

### 2020-04-06 draw.rb

@@ -3,7 +3,7 @@ data = []
 File.open(ARGV[0]).read.each_line do |l|
   ncss, flag = l.strip.split(',', 2)
   ncss = ncss.to_i
-  next if ncss > 2000
+  next if ncss > 600
   data << { ncss: ncss, immutable: flag == 'yes' }
 end
 
@@ -23,16 +23,21 @@ cols = []
   }
 end
 
+max = cols.map { |c| c[:total] }.max
+
+puts "% Max: #{max}"
 puts "% Lo: #{lo}"
 puts "% Hi: #{hi}"
 puts "% Step: #{step} (steps: #{steps})"
 
-puts '\begin{tikzpicture}' +
-'\begin{axis}[axis lines=middle, xlabel={NCSS}, ylabel={$n_i$},' +
-"xmin=#{lo}, xmax=#{hi}, ymin=0, ymax=1," +
-"extra tick style={major grid style=black}," +
-'grid=major]\addplot [only marks, mark size=1pt] table {' + "\n"
+puts '\begin{tikzpicture}'
+puts '\begin{axis}[width=12cm,height=8cm,'
+puts 'axis lines=middle, xlabel={NCSS}, ylabel={},'
+puts "xmin=#{lo}, xmax=#{hi}, ymin=0, ymax=100,"
+puts 'yticklabel={\pgfmathprintnumber{\tick}\%},'
+puts "extra tick style={major grid style=black},grid=major]"
 cols.each_with_index do |v, i|
-  puts "  #{lo + i * step} #{v[:share]}\n"
+  puts "\\addplot [only marks, mark size=#{[8 * v[:total] / max,0.5].max}pt] coordinates {"
+  puts "(#{lo + i * step},#{v[:share] * 100})};"
 end
-puts '};\end{axis}\end{tikzpicture}'
+puts '\end{axis}\end{tikzpicture}'

### 2020-04-06 paper1/article.tex
more

@@
 \bibliographystyle{plainurl}
-\title{The Impact of Class Immutability on NCSS}
-\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\Moscow, Russia\\yegor.bugayenko@huawei.com}{}{}
+\title{The Impact of Immutability on Class Size}
+\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\Moscow, Russia\\\texttt{yegor.bugayenko@huawei.com}}{}{}
 % \authorrunning{Yegor Bugayenko}
@@
 \begin{abstract}
-...
+To be written later...
 \end{abstract}
@@
 The list included popular Java open source products, like
-Spring Framework, RxJava, Google Guava, and others.
+Spring, RxJava, Guava, MyBatis, Clojure, JUnit, Lombok,
+Graal, Selenium, Spark, Mockito, Jenkins, Netty, and others.
 

### 2020-04-06 draw.rb

@@ -3,7 +3,7 @@ data = []
 File.open(ARGV[0]).read.each_line do |l|
   ncss, flag = l.strip.split(',', 2)
   ncss = ncss.to_i
-  next if ncss > 600
+  next if ncss > 1000
   data << { ncss: ncss, immutable: flag == 'yes' }
 end
 
@@ -31,7 +31,7 @@ puts "% Hi: #{hi}"
 puts "% Step: #{step} (steps: #{steps})"
 
 puts '\begin{tikzpicture}'
-puts '\begin{axis}[width=12cm,height=8cm,'
+puts '\begin{axis}[width=14cm,height=10cm,'
 puts 'axis lines=middle, xlabel={NCSS}, ylabel={},'
 puts "xmin=#{lo}, xmax=#{hi}, ymin=0, ymax=100,"
 puts 'yticklabel={\pgfmathprintnumber{\tick}\%},'

### 2020-04-06 draw.rb
more

@@ -1,19 +1,28 @@
+#!/usr/bin/env ruby
+
+require 'slop'
+
+opts = Slop.parse(ARGV, strict: true, help: true) do |o|
+  o.integer '--max-ncss', default: 1000
+  o.integer '--steps', default: 100
+  o.integer '--circle-size', default: 8
+  o.string '--summary', default: 'summary.csv'
+end
 
 data = []
-File.open(ARGV[0]).read.each_line do |l|
+File.open(opts[:summary]).read.each_line do |l|
   ncss, flag = l.strip.split(',', 2)
   ncss = ncss.to_i
-  next if ncss > 1000
+  next if ncss > opts['max-ncss']
   data << { ncss: ncss, immutable: flag == 'yes' }
 end
 
-steps = 100
 lo = data.map { |r| r[:ncss] }.min
 hi = data.map { |r| r[:ncss] }.max
-step = (hi - lo).to_f / steps
+step = (hi - lo).to_f / opts[:steps]
 
 cols = []
-(0..steps - 1).each do |s|
+(0..opts[:steps] - 1).each do |s|
   all = data.select { |r| r[:ncss] > s * step && r[:ncss] < s * (step + 1) }
   total = all.count
   immutable = all.select { |r| r[:immutable] }.count
@@ -28,7 +37,7 @@ max = cols.map { |c| c[:total] }.max
 puts "% Max: #{max}"
 puts "% Lo: #{lo}"
 puts "% Hi: #{hi}"
-puts "% Step: #{step} (steps: #{steps})"
+puts "% Step: #{step} (steps: #{opts[:steps]})"
 
 puts '\begin{tikzpicture}'
 puts '\begin{axis}[width=14cm,height=10cm,'
@@ -37,7 +46,7 @@ puts "xmin=#{lo}, xmax=#{hi}, ymin=0, ymax=100,"
 puts 'yticklabel={\pgfmathprintnumber{\tick}\%},'
 puts "extra tick style={major grid style=black},grid=major]"
 cols.each_with_index do |v, i|
-  puts "\\addplot [only marks, mark size=#{[8 * v[:total] / max,0.5].max}pt] coordinates {"
+  puts "\\addplot [only marks, mark size=#{[opts['circle-size'] * v[:total] / max,0.5].max}pt] coordinates {"
   puts "(#{lo + i * step},#{v[:share] * 100})};"
 end
 puts '\end{axis}\end{tikzpicture}'

### 2020-04-06 paper/article.tex
rename

--- 2020-04-06 paper1/article.tex
@@
 all classes with the given NCSS. The diameter of the plot on the graph
-is related to the amount of classes found for the specific NCSS.
+is related to the amount of classes found for the specific NCSS. Classes
+with NCSS larger than 1000 are excluded from the graph for the sake of
+visual elegancy.
 
@@
 
+The Figure~\ref{fig:2} demonstrates classes with NCSS lower than 200.
+
+\begin{figure}[h]
+  \input{../ncss-closer.tex}
+  \caption{Closer look to the lower part of the distribution}
+  \label{fig:2}
+\end{figure}
+
 \section{Conclusion}
@@
 
+The source code of Ruby and Python scripts used to do the research
+is available in GitHub repository \texttt{yegor256/size-vs-immutability}.
+
 \end{document}

### 2020-04-07 paper/article.tex
bib

--- 2020-04-06 paper1/article.tex
@@
 \title{The Impact of Immutability on Class Size}
-\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\Moscow, Russia\\\texttt{yegor.bugayenko@huawei.com}}{}{}
+\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\\texttt{yegor.bugayenko@huawei.com}}{}{}
 % \authorrunning{Yegor Bugayenko}
@@
 to them.
+Immutable classes greatly simplify programming, program maintenance, and
+reasoning about programs. Immutable classes can be freely shared, even between
+concurrent threads and with untrusted code, without the need to worry
+about modifications, even temporary ones, that could result in
+inconsistent states or broken invariants.
+Immutability is a recommended coding practice for Java~\cite{bloch2016}.
 
@@
 Spring, RxJava, Guava, MyBatis, Clojure, JUnit, Lombok,
-Graal, Selenium, Spark, Mockito, Jenkins, Netty, and others.
+Graal, Selenium, Spark, Mockito, Neo4j, Jenkins, Netty, and others.
 
@@
 There were \totaljavafiles{} files found. Classes without a single
-non-\texttt{static} attribute were excluded (including interfaces and enums).
+non-\texttt{static} attribute were excluded (such as utility classes,
+interfaces, or enums).
 
@@
 A class was considered immutable if it didn't have any
-attributes without \texttt{final} modified.
+attributes without \texttt{final} modifier.
 
 The Figure~\ref{fig:1} demonstrates the results obtained. The x-axis is
-the value of NCSS. The y-axis is the share of immutable classes among
+the value of NCSS of Java classes. The y-axis is the share of immutable classes among
 all classes with the given NCSS. The diameter of the plot on the graph
-is related to the amount of classes found for the specific NCSS.
+is related to the amount of classes found for the specific NCSS. Classes
+with NCSS larger than 1000 were excluded from the graph for the sake of
+visual elegancy.
 
@@
 
+\section{Discussion}
+
+The notion of immutability is not as straightforward as it might seem, and many
+different definitions of immutability exist~\cite{potanin2013}.
+
 \section{Conclusion}
@@
 
+The source code of Ruby and Python scripts used to do the research
+is available in GitHub repository \texttt{yegor256/size-vs-immutability}.
+
+\bibliography{main}
+
 \end{document}

### 2020-04-07 calc.py

@@ -33,10 +33,11 @@ def ncss(tree):
       metric += 1
   return metric
 
-with open(sys.argv[1], encoding='utf-8') as f:
+java = sys.argv[1]
+with open(java, encoding='utf-8') as f:
   try:
     raw = javalang.parse.parse(f.read())
     tree = raw.filter(javalang.tree.ClassDeclaration)
-    print(str(ncss(raw)) + ',' + ('yes' if immutable(tree) else 'no'))
+    print(str(ncss(raw)) + ',' + ('yes' if immutable(tree) else 'no') + ',' + java)
   except Exception as e:
-    sys.exit(e.args)
+    sys.exit(str(e) + ': ' + java)

### 2020-04-07 draw.rb

@@ -11,7 +11,7 @@ end
 
 data = []
 File.open(opts[:summary]).read.each_line do |l|
-  ncss, flag = l.strip.split(',', 2)
+  ncss, flag = l.strip.split(',')
   ncss = ncss.to_i
   next if ncss > opts['max-ncss']
   data << { ncss: ncss, immutable: flag == 'yes' }

### 2020-04-07 draw.rb
more

@@ -3,6 +3,8 @@
 require 'slop'
 
 opts = Slop.parse(ARGV, strict: true, help: true) do |o|
+  o.integer '--width', default: 12
+  o.integer '--height', default: 8
   o.integer '--max-ncss', default: 1000
   o.integer '--steps', default: 100
   o.integer '--circle-size', default: 8
@@ -40,13 +42,13 @@ puts "% Hi: #{hi}"
 puts "% Step: #{step} (steps: #{opts[:steps]})"
 
 puts '\begin{tikzpicture}'
-puts '\begin{axis}[width=14cm,height=10cm,'
+puts "\\begin{axis}[width=#{opts[:width]}cm,height=#{opts[:height]}cm,"
 puts 'axis lines=middle, xlabel={NCSS}, ylabel={},'
 puts "xmin=#{lo}, xmax=#{hi}, ymin=0, ymax=100,"
 puts 'yticklabel={\pgfmathprintnumber{\tick}\%},'
 puts "extra tick style={major grid style=black},grid=major]"
 cols.each_with_index do |v, i|
-  puts "\\addplot [only marks, mark size=#{[opts['circle-size'] * v[:total] / max,0.5].max}pt] coordinates {"
+  puts "\\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=#{[opts['circle-size'] * v[:total] / max,0.5].max}pt] coordinates {"
   puts "(#{lo + i * step},#{v[:share] * 100})};"
 end
 puts '\end{axis}\end{tikzpicture}'

### 2020-04-09 paper/article.tex
github action

@@
 
-No similar research has been done so far.
+Object immutability has been the subject of multiple researches since
+the begining of object-oriented programming. Even though, according to~\cite{potanin2013},
+``the notion of immutability is not as straightforward as it might seem, and many
+different definitions of immutability exist,'' ...
+
+Even though Java encourages programmers to explicitly make objects immutable
+by attaching the \texttt{final} modifier to its attributes, very often, as was
+demonstrated by~\cite{unkel2008}, the modifier is not used even when
+attributes are not modified after object instantiation or, as demonstrated
+by~\cite{nelson2012}, after initialization. Such a delayed initialization of attributes, as
+explained by~\cite{fahndrich2007}, leads to the prevalence of \texttt{null}
+and the ability to initialize circular data structures, while both of these
+practices have negative impact on software quality.
+
+Despite the existence of the \texttt{final} modifier, Reflection API enables
+modification of any attributes in Java. Some mechanisms were suggested to
+overcome this technical vulnerability, for example
+via ``freezer'' objects~\cite{leino2008},
+a new type system~\cite{tschantz2005},
+or immutability assertion framework~\cite{pechtchanski2005}.
+However, without any modifications to Java~\cite{steele2005},
+thanks to the existence of Reflection API,
+it's not possible to say whether a particular Java object is modified
+after instantiation or not---any object can be modified.
+
+Aside from the \texttt{final} modifier, which is an explicit way of declaring
+read-only status of an attribute, there are also methods of
+code analysis, enabling the detection of implicit immutability. For example,
+a static flow-sensitive analysis algorithm was introduced by~\cite{porat2000}
+to classify fields and classes as either mutable or immutable.
 
@@
 
-The notion of immutability is not as straightforward as it might seem, and many
-different definitions of immutability exist~\cite{potanin2013}.
+...
 

### 2020-04-09 paper/ncss.tex
summaries

% Max: 3311
% Lo: 2
% Hi: 1000
% Step: 9.98 (steps: 100)
\begin{tikzpicture}
\begin{axis}[width=12cm,height=8cm,
axis lines=middle, xlabel={NCSS}, ylabel={},
xmin=2, xmax=1000, ymin=0, ymax=100,
yticklabel={\pgfmathprintnumber{\tick}\%},
extra tick style={major grid style=black},grid=major]
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(2.0,0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(11.98,51.86260280599904)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=7pt] coordinates {
(21.96,40.19230769230769)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=7pt] coordinates {
(31.94,40.29714978775015)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=8pt] coordinates {
(41.92,36.24282694050136)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=7pt] coordinates {
(51.900000000000006,32.57948213700426)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=6pt] coordinates {
(61.88,30.265995686556437)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=6pt] coordinates {
(71.86,28.994082840236686)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=5pt] coordinates {
(81.84,27.139874739039666)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=5pt] coordinates {
(91.82000000000001,25.547445255474454)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(101.80000000000001,24.345679012345677)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(111.78,24.395604395604394)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=4pt] coordinates {
(121.76,21.356502242152466)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(131.74,18.953846153846154)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(141.72,19.180887372013654)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(151.70000000000002,19.58997722095672)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(161.68,16.338880484114977)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=3pt] coordinates {
(171.66,15.89825119236884)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(181.64000000000001,15.646258503401361)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(191.62,16.25794732061762)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(201.60000000000002,17.51412429378531)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(211.58,16.151545363908276)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(221.56,13.934426229508196)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(231.54000000000002,14.146868250539956)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=2pt] coordinates {
(241.52,14.387633769322235)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(251.5,15.590742996345918)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(261.48,15.031847133757962)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(271.46000000000004,14.701803051317613)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(281.44,12.949640287769784)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(291.42,12.669683257918551)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(301.40000000000003,9.923664122137405)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(311.38,9.748427672955975)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(321.36,9.317803660565724)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(331.34000000000003,10.35120147874307)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(341.32,9.454545454545455)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(351.3,10.054844606946983)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(361.28000000000003,9.496124031007753)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(371.26,9.421841541755889)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(381.24,10.199556541019955)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(391.22,9.647058823529411)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=1pt] coordinates {
(401.20000000000005,9.156626506024097)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(411.18,9.135802469135802)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(421.16,7.567567567567568)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(431.14000000000004,8.743169398907105)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(441.12,9.63855421686747)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(451.1,9.433962264150944)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(461.08000000000004,9.67741935483871)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(471.06,9.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(481.04,8.411214953271028)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(491.02000000000004,7.006369426751593)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(501.0,6.666666666666667)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(510.98,7.241379310344828)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(520.96,6.474820143884892)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(530.94,6.172839506172839)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(540.9200000000001,7.456140350877193)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(550.9,8.071748878923767)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(560.88,6.422018348623854)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(570.86,9.251101321585903)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(580.84,7.456140350877193)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(590.82,8.81057268722467)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(600.8000000000001,8.695652173913043)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(610.78,9.813084112149532)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(620.76,9.047619047619047)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(630.74,8.653846153846153)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(640.72,10.65989847715736)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(650.7,9.473684210526317)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(660.6800000000001,10.05586592178771)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(670.6600000000001,9.35672514619883)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(680.64,10.828025477707007)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(690.62,10.738255033557047)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(700.6,9.45945945945946)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(710.58,10.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(720.5600000000001,8.823529411764707)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(730.5400000000001,6.8181818181818175)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(740.52,5.147058823529411)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(750.5,4.794520547945205)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(760.48,4.081632653061225)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(770.46,3.64963503649635)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(780.44,4.166666666666666)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(790.4200000000001,4.25531914893617)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(800.4000000000001,4.316546762589928)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(810.38,4.477611940298507)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(820.36,5.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(830.34,3.8095238095238098)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(840.32,6.0606060606060606)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(850.3000000000001,5.208333333333334)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(860.2800000000001,5.747126436781609)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(870.26,4.761904761904762)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(880.24,3.7037037037037033)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(890.22,3.614457831325301)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(900.2,3.7037037037037033)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(910.1800000000001,3.571428571428571)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(920.1600000000001,1.25)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(930.14,1.4084507042253522)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(940.12,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(950.1,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(960.08,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(970.0600000000001,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(980.0400000000001,0.0)};
\addplot [only marks, mark=*, mark options={fill=lightgray,draw=black}, mark size=0.5pt] coordinates {
(990.0200000000001,0.0)};
\end{axis}\end{tikzpicture}

### 2020-04-10 paper/article.tex
summaries
deep immutability

@@
   {\medskip}
-\input{../total}
+\input{total}
 \begin{document}
@@
 it's not possible to say whether a particular Java object is modified
-after instantiation or not---any object can be modified.
+after instantiation or not---any object can be modified. As was noticed
+by~\cite{hakonen1999} in relation to immutability,
+``none of the current OO languages can prevent the programmer from writing
+a piece of code which violates the integrity of an object''
 
@@
 to classify fields and classes as either mutable or immutable.
+
+Even if programmers do not violate encapsulation via reflection, the
+modified \texttt{final} only guarantees shallow immutability, as
+explained by~\cite{hakonen1999}. Deep immutability, on the other hand,
+exists only if the object pointed by the attributes are deeply immutable
+(recursively).
 
@@
 \begin{figure}[h]
-  \input{../ncss.tex}
+  \input{ncss.tex}
   \caption{Distribution of class immutability by NCSS}

### 2020-04-10 paper/article.tex
more

--- 2020-04-07 paper/article.tex
@@
 \documentclass[12pt]{article}
-\bibliographystyle{plainurl}
 \title{The Impact of Immutability on Class Size}
@@
 \usepackage[utf8]{inputenc}
+\usepackage[numbers]{natbib}
+\bibliographystyle{plainnat}
 \usepackage[inline]{enumitem}
@@
 \begin{abstract}
-To be written later...
+According to subjective opinions of many industry experts,
+object immutability is a virtue in object-oriented programming, since it
+leads to side-effect free design, cleaner code, better concurrency,
+and many other factors. However, it has never been empirically
+demonstrated exactly how immutability affects quality metrics of
+object-oriented program. In the following research we analyzed
+\thetotaljavafiles{} classes from \thetotalrepos{} public Java repositories
+to find out how immutability affects the size of code.
 \end{abstract}
@@
 inconsistent states or broken invariants.
-Immutability is a recommended coding practice for Java~\cite{bloch2016}.
+Immutability is a recommended coding practice for Java~\citep{bloch2016}.
 
@@
 
-No similar research has been done so far.
+Object immutability has been the subject of multiple researches since
+the begining of object-oriented programming~\citep{hakonen1999,porat2000,lea2000}.
+Even though, according to~\citet{potanin2013},
+``the notion of immutability is not as straightforward as it might seem, and many
+different definitions of immutability exist,'' most industry experts consider
+immutability being a virtue of classes and objects in Java and other
+object-oriented programming languages.
+
+Even though Java encourages programmers to explicitly make objects immutable
+by attaching the \texttt{final} modifier to its attributes, very often, as was
+demonstrated by~\citet{unkel2008}, the modifier is not used even when
+attributes are not modified after object instantiation or, as demonstrated
+by~\citet{nelson2012}, after initialization. Such a delayed initialization of attributes, as
+explained by~\citet{fahndrich2007}, leads to the prevalence of \texttt{null}
+and the ability to initialize circular data structures, while both of these
+practices have negative impact on software quality.
+
+Despite the existence of the \texttt{final} modifier, Reflection API enables
+modification of any attributes in Java. Some mechanisms were suggested to
+overcome this technical vulnerability, for example
+via ``freezer'' objects~\citep{leino2008},
+a new type system~\citep{tschantz2005},
+or immutability assertion framework~\citep{pechtchanski2005}.
+However, without any modifications to Java~\citep{steele2005},
+thanks to the existence of Reflection API,
+it's not possible to say whether a particular Java object is modified
+after instantiation or not---any object can be modified. As was noticed
+by~\citet{hakonen1999} in relation to immutability,
+``none of the current OO languages can prevent the programmer from writing
+a piece of code which violates the integrity of an object''
+
+Aside from the \texttt{final} modifier, which is an explicit way of declaring
+read-only status of an attribute, there are also methods of
+code analysis, enabling the detection of implicit immutability. For example,
+a static flow-sensitive analysis algorithm was introduced by~\citet{porat2000}
+to classify fields and classes as either mutable or immutable.
+
+Even if programmers do not violate encapsulation via reflection, the
+modified \texttt{final} only guarantees shallow immutability, as
+explained by~\citet{hakonen1999}. Deep immutability, on the other hand,
+exists only if the object pointed by the attributes are deeply immutable
+(recursively).
+
+To our knowledge, the impact of immutability of class size has not been
+analyzed yet.
 
@@
 A list of Java repositories were retrieved from GitHub via their
-public API. The first \totalrepos{} repositories were taken, which satisfied
+public API. The first \thetotalrepos{} repositories were taken, which satisfied
 the selection criteria:
 \begin{enumerate*}[label={\arabic*)}]
-\item more than 1000 stars,
+\item more than 1,000 GitHub stars,
 \item more than 200 Kb of data,
-\item not archived,
+\item not archived, and
 \item public.
 \end{enumerate*}
-The list included popular Java open source products, like
+The list included popular Java open source products, such as
 Spring, RxJava, Guava, MyBatis, Clojure, JUnit, Lombok,
@@
 Files with \texttt{.java} extension were taken from all repositories.
-There were \totaljavafiles{} files found. Classes without a single
+There were \thetotaljavafiles{} files found. Classes without a single
 non-\texttt{static} attribute were excluded (such as utility classes,
-interfaces, or enums).
+interfaces, or enums), despite the fact that some OOP
+experts consider such classes valid and immutable, saying that
+``the simplest immutable objects have no internal fields at all''~\citep{lea2000}.
+We decided to not take these classes into account because they do not instantiate objects
+and because of that do not belong to object-oriented paradigm,
+as explained by~\cite{west2004,bugayenko2017}.
 
-NCSS metric and immutability were calculated for each Java class.
+NCSS metric and immutability were calculated for each Java class,
+using javalang, Java-parsing library written in Python.
 A class was considered immutable if it didn't have any
@@
 visual elegancy.
+The right side of the graph has mostly (over 90\%) mutable classes. The largest
+value of NCSS observed was \thelargestncss{}.
 
 \begin{figure}[h]
-  \input{../ncss.tex}
+  \input{ncss.tex}
   \caption{Distribution of class immutability by NCSS}
@@
 
-\section{Discussion}
-
-The notion of immutability is not as straightforward as it might seem, and many
-different definitions of immutability exist~\cite{potanin2013}.
-
 \section{Conclusion}
@@
 It was empirically confirmed that larger Java classes tend to be immutable
-less frequently than smaller ones.
+less frequently than smaller ones. At the same time it is possible to conclude
+that immutable classes tend to be smaller. Both conclusions justify
+the hypothesis that immutability in object-oriented programming leads
+to higher maintainability, since classes will be smaller.
 

### 2020-04-11 paper/article.tex
address off

@@
 \title{The Impact of Immutability on Class Size}
-\author{Yegor Bugayenko\\Huawei Technologies Co., Ltd.\\Russian Research Institute (RRI)\\\texttt{yegor.bugayenko@huawei.com}}{}{}
+\author{Yegor Bugayenko}{}{}
 % \authorrunning{Yegor Bugayenko}
@@
 Object immutability has been the subject of multiple researches since
-the begining of object-oriented programming~\citep{hakonen1999,porat2000,lea2000}.
+the begining of object-oriented programming~\citep{hakonen1999,porat2000,lea2000,taivalsaari1993}.
 Even though, according to~\citet{potanin2013},

### 2020-04-12 paper/article.tex
badge

@@
 
-One of the most important factors negatively affecting the quality
-of object-oriented software is the length of the code. Larger classes
-cause readability and maintability issues. One of the simplest way
-to calculate size of a Java class is via the NCSS metric, which
+On one hand, one of the most important factors negatively affecting the quality
+of object-oriented software is the \emph{size} of classes, as was
+demonstrated by~\citet{li1993,al2013}: classes with lower size have better maintainability.
+One of the simplest way
+to calculate the size of a Java class is via the NCSS metric, which
 stands for Non-Commenting Source Statements. Larger classes with bigger
@@
 
-Immutability is a property of a class in OOP. A class is immutable if
+On the other hand, \emph{immutability} is a property of a class in OOP. A class is immutable if
 it's impossible to modify its attributes after instantiation. In Java this
@@
 This research empirically validates the relationship between immutability
-and NCSS.
+and class size.
 
@@
 immutability being a virtue of classes and objects in Java and other
-object-oriented programming languages.
+object-oriented programming languages~\citep{bloch2016}. There a few important
+concerns of immutability addresses so far by object-oriented researchers.
 
-Even though Java encourages programmers to explicitly make objects immutable
+First, even though Java encourages programmers to explicitly make objects immutable
 by attaching the \texttt{final} modifier to its attributes, very often, as was
@@
 
-Despite the existence of the \texttt{final} modifier, Reflection API enables
+Second, despite the existence of the \texttt{final} modifier, Reflection API enables
 modification of any attributes in Java. Some mechanisms were suggested to
@@
 
-Aside from the \texttt{final} modifier, which is an explicit way of declaring
+Third, aside from the \texttt{final} modifier, which is an explicit way of declaring
 read-only status of an attribute, there are also methods of
@@
 
-Even if programmers do not violate encapsulation via reflection, the
+Forth, even if programmers do not violate encapsulation via reflection, the
 modified \texttt{final} only guarantees shallow immutability, as
@@
 
-To our knowledge, the impact of immutability of class size has not been
+However, to our knowledge, the impact of immutability of class size has not been
 analyzed yet.
@@
 NCSS metric and immutability were calculated for each Java class,
-using javalang, Java-parsing library written in Python.
+using \texttt{javalang}, an open source Java-parsing library written in Python.
 A class was considered immutable if it didn't have any
@@
 is related to the amount of classes found for the specific NCSS. Classes
-with NCSS larger than 1000 were excluded from the graph for the sake of
-visual elegancy.
+with NCSS larger than 1000 were excluded from the graph. This decision was
+motivated by the confounding effect the size of a class has
+on the validity of object-oriented metrics, as was empirically shown by~\citet{el2001}.
 The right side of the graph has mostly (over 90\%) mutable classes. The largest
@@
 The tendency is visually obvious: larger classes are less often immutable. It
-is also obvious that the majority of classes have NCSS values smaller than 100.
+was also observed that the majority of classes have NCSS values smaller than 100:
+\thebelowhundred{} classes out of \thetotaljavafiles{} total.
 

### 2020-04-12 calc.py
MIT license

@@ -1,3 +1,25 @@
+# The MIT License (MIT)
+#
+# Copyright (c) 2020 Yegor Bugayenko
+#
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+#
+# The above copyright notice and this permission notice shall be included
+# in all copies or substantial portions of the Software.
+#
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
+
 import sys
 import javalang
 

### 2020-04-12 draw.rb

@@ -1,4 +1,25 @@
 #!/usr/bin/env ruby
+# The MIT License (MIT)
+#
+# Copyright (c) 2020 Yegor Bugayenko
+#
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+#
+# The above copyright notice and this permission notice shall be included
+# in all copies or substantial portions of the Software.
+#
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
 
 require 'slop'
 

### 2020-04-12 find-repos.rb

@@ -1,4 +1,25 @@
 #!/usr/bin/env ruby
+# The MIT License (MIT)
+#
+# Copyright (c) 2020 Yegor Bugayenko
+#
+# Permission is hereby granted, free of charge, to any person obtaining a copy
+# of this software and associated documentation files (the "Software"), to deal
+# in the Software without restriction, including without limitation the rights
+# to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+# copies of the Software, and to permit persons to whom the Software is
+# furnished to do so, subject to the following conditions:
+#
+# The above copyright notice and this permission notice shall be included
+# in all copies or substantial portions of the Software.
+#
+# THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+# IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+# FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
+# AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+# LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+# OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+# SOFTWARE.
 
 STDOUT.sync = true
 

### 2020-04-19 paper/article.tex
MIT license
more

@@
-% \documentclass[a4paper,UKenglish,cleveref, autoref]{lipics-v2019}
-\documentclass[12pt]{article}
-\title{The Impact of Immutability on Class Size}
+% The MIT License (MIT)
+%
+% Copyright (c) 2020 Yegor Bugayenko
+%
+% Permission is hereby granted, free of charge, to any person obtaining a copy
+% of this software and associated documentation files (the "Software"), to deal
+% in the Software without restriction, including without limitation the rights
+% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
+% copies of the Software, and to permit persons to whom the Software is
+% furnished to do so, subject to the following conditions:
+%
+% The above copyright notice and this permission notice shall be included
+% in all copies or substantial portions of the Software.
+%
+% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
+% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
+% FITNESS FOR A PARTICULAR PURPOSE AND NON-INFRINGEMENT. IN NO EVENT SHALL THE
+% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
+% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
+% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE
+% SOFTWARE.
+
+\documentclass[sigconf]{acmart}
+\title{The Impact of Object Immutability on the Java Class Size}
 \author{Yegor Bugayenko}{}{}
-% \authorrunning{Yegor Bugayenko}
-% \ccsdesc[100]{Software and its engineering~Object oriented languages}
-% \keywords{Object-Oriented Programming, Immutability, NCSS}
+\email{yegor.bugayenko@huawei.com}
+\affiliation{%
+  \institution{Huawei Technologies Co., Ltd.}
+  \city{Moscow, Russia}
+}
+\ccsdesc[100]{Software and its engineering~Object oriented languages}
+\keywords{Object-Oriented Programming, Immutability, NCSS}
+\setcopyright{acmcopyright}
+\copyrightyear{2020}
+\acmYear{2020}
+
 \usepackage[utf8]{inputenc}
-\usepackage[numbers]{natbib}
-\bibliographystyle{plainnat}
 \usepackage[inline]{enumitem}
@@
 \begin{document}
-\raggedbottom
-\maketitle
 
@@
 \end{abstract}
+\maketitle
 
@@
 
+The paper is organized as follows.
+Section~\ref{sec:background} defines various terms used in the paper.
+Section~\ref{sec:related} covers related work in the areas of immutability and class size.
+Section~\ref{sec:results} covers our empirical case study of \thetotalrepos{} Java source code repositories.
+Section~\ref{sec:discussion} covers limitations of both the metric and the study.
+Finally, we summarize our study in Section~\ref{sec:conclusion}.
+
+\section{Background}
+\label{sec:background}
+
+Immutability in Java is a compile-time restriction of an object, making it
+impossible for other objects to modify any of its attributes. For example,
+objects of this class are immutable, thanks to the \texttt{final} modifier
+attached to all of its attributes:
+
+\begin{minted}{text}
+class Book {
+  private final int id;
+  private final String title;
+  Book(int i, String t) {
+    this.id = i;
+    this.title = t;
+  }
+}
+\end{minted}
+
+The only way to modify \texttt{final} attributes is through the constructor
+of the class, which in Java by convention has the same name as the class (\text{Book}
+in the example above).
+
+A class of any size may be immutable. A class with no attributes is immutable.
+A utility class, which is a class with no attributes, many static methods,
+and a private constructor, is immutable.
+
 \section{Related Work}
+\label{sec:related}
 
@@
 
-However, to our knowledge, the impact of immutability of class size has not been
-analyzed yet.
+However, to our knowledge, the impact of immutability on the Java
+class size has not been analyzed yet.
 
 \section{Empirical Results}
+\label{sec:results}
 
@@
 
+\section{Discussion}
+\label{sec:discussion}
+
+To be written...
+
 \section{Conclusion}
+\label{sec:conclusion}
 
@@
 
-The source code of Ruby and Python scripts used to do the research
-is available in GitHub repository \texttt{yegor256/size-vs-immutability}.
+% The source code of Ruby and Python scripts used to do the research
+% is available in GitHub repository \texttt{yegor256/size-vs-immutability}.
 
+\bibliographystyle{ACM-Reference-Format}
 \bibliography{main}

### 2020-04-19 draw.rb

@@ -24,11 +24,11 @@
 require 'slop'
 
 opts = Slop.parse(ARGV, strict: true, help: true) do |o|
-  o.integer '--width', default: 12
-  o.integer '--height', default: 8
+  o.integer '--width', default: 9
+  o.integer '--height', default: 5
   o.integer '--max-ncss', default: 1000
   o.integer '--steps', default: 100
-  o.integer '--circle-size', default: 8
+  o.integer '--circle-size', default: 6
   o.string '--summary', default: 'summary.csv'
 end
 

### 2020-06-02 paper/article.tex
new format

@@
 
-\documentclass[sigconf]{acmart}
+\documentclass[3p,times,procedia]{elsarticle}
+\usepackage{ecrc}
+\volume{00}
+\firstpage{1}
+\journalname{Procedia Computer Science}
+\runauth{Yegor Bugayenko et al.}
+\jid{procs}
 \title{The Impact of Object Immutability on the Java Class Size}
-\author{Yegor Bugayenko}{}{}
-\email{yegor.bugayenko@huawei.com}
-\affiliation{%
-  \institution{Huawei Technologies Co., Ltd.}
-  \city{Moscow, Russia}
-}
-\ccsdesc[100]{Software and its engineering~Object oriented languages}
-\keywords{Object-Oriented Programming, Immutability, NCSS}
-\setcopyright{acmcopyright}
-\copyrightyear{2020}
-\acmYear{2020}
 
+\usepackage[bookmarks=false]{hyperref}
+  \hypersetup{colorlinks,linkcolor=blue,citecolor=blue,urlcolor=blue}
 \usepackage[utf8]{inputenc}
@@
 \begin{document}
+\begin{frontmatter}
 
+\dochead{24th International Conference on Knowledge-Based and Intelligent Information \& Engineering Systems}%
+\title{The Impact of Object Immutability on the Java Class Size}
+\author[a]{Yegor Bugayenko}
+\author[b]{Sergey Zykov}
+\address[a]{Huawei Technologies Co., Ltd., Moscow, Russia}
+\address[b]{Higher School of Economics, Moscow, Russia}
 \begin{abstract}
-According to subjective opinions of many industry experts,
+According to the subjective opinions of many industry experts,
 object immutability is a virtue in object-oriented programming, since it
-leads to side-effect free design, cleaner code, better concurrency,
+leads to side-effect-free design, cleaner code, better concurrency,
 and many other factors. However, it has never been empirically
 demonstrated exactly how immutability affects quality metrics of
-object-oriented program. In the following research we analyzed
+object-oriented programs. In the following research, we analyzed
 \thetotaljavafiles{} classes from \thetotalrepos{} public Java repositories
-to find out how immutability affects the size of code.
+to find out how immutability affects the size of the code.
 \end{abstract}
-\maketitle
+\begin{keyword}Object-Oriented Programming; Immutability; NCSS\end{keyword}
+\end{frontmatter}
 
@@
 The only way to modify \texttt{final} attributes is through the constructor
-of the class, which in Java by convention has the same name as the class (\text{Book}
+of the class, which in Java by convention has the same name as the class (\texttt{Book}
 in the example above).
@@
 
-To be written...
+The results obtained empirically confirm the hyposesis that immutability
+not only leads to better concurrency, side-effect free design, cleaner
+contracts between classes, but also to smaller classes, by the amount of
+methods and attributes inside.
+
+This is a practical comparison example of two Java libraries designed
+by two different groups of programmers for the same purpose: sending emails from Java code.
+The first one is \code{commons-email} (version 1.5) by Apache with
+a large mutable class \code{SimpleEmail} at the core.
+The second one is \code{jcabi-email} (version 1.10) with a set of immutable classes.
+Here is how Java source code may look, if it sends an email using \code{commons-email}:
+
+\begin{minted}{text}
+Email email = new SimpleEmail();
+email.setHostName("smtp.googlemail.com");
+email.setSmtpPort(465);
+email.setAuthenticator(new DefaultAuthenticator("user", "pwd"));
+email.setFrom("yegor256@gmail.com", "Yegor Bugayenko");
+email.addTo("friend@jcabi.com");
+email.setSubject("How are you?");
+email.setMsg("Hi, how are you?");
+email.send();
+\end{minted}
+
+Here is how the same email sending scenario would be
+implemented with \code{jcabi-email}:
+
+\begin{minted}{text}
+Postman postman = new Postman.Default(
+  new SMTP("smtp.googlemail.com", 465, "user", "pwd")
+);
+Envelope envelope = new Envelope.MIME(
+  new Array<Stamp>(
+    new StSender("Yegor Bugayenko <yegor256@gmail.com>"),
+    new StRecipient("friend@jcabi.com"),
+    new StSubject("How are you?")
+  ),
+  new Array<Enclosure>(
+    new EnPlain("Hi, how are you?")
+  )
+);
+postman.send(envelope);
+\end{minted}
+
+In the first example, it is a monster \code{SimpleEmail} class that can do many things, including
+sending MIME message via SMTP, creating the message, configuring
+its parameters, adding MIME parts to it, and so on. There are 33 private properties, over
+a hundred methods, and about two thousands lines of code.
+
+In the second example, there are seven objects instantiated via seven \code{new} calls.
+\code{Postman} is responsible for packaging a MIME message; \code{SMTP} is
+responsible for sending it via SMTP; stamps (\code{StSender}, \code{StRecipient},
+and \code{StSubject}) are responsible for configuring the MIME message before
+delivery; enclosure \code{EnPlain} is responsible for creating a MIME part
+for the message, which is going to be send. These seven objects being constructed,
+encapsulating one into another, and then the postman is asked to \code{send()}
+the envelope over the wire.
+
+From a user perspective, there is almost nothing wrong. \code{SimpleEmail} is a powerful
+class with multiple controls---a project needs to pick the right one and the job
+gets done. However, from a developer perspective \code{SimpleEmail} class is
+very difficult to maintain, mostly because the class is large.
+Multiple getters and setters, which are the control points of the class, modify
+object attributes, configuring its behavior. When a new functionality is
+required, a developer has to add new attributes to the class and a new
+pair of setters and getters. Of course, such a modification decreases the
+cohesion of the class, since there is very little or no interconnection
+between newly added attributes and previously existing ones. Every new method
+added to such a big class, turns into an isolated island of functionality,
+with its own set of attributes.
+
+The immutability of a class makes it difficult to make a class larger
+without spending a big amount of efforts for refactoring.
+If the \code{SimpleEmail} class would be immutable
+in the beginning, it wouldn't be possible to add so many methods into it
+and encapsulate so many properties. Because an immutable object only accepts a state
+through its constructors. It's difficult to imagine a 33-argument constructor.
+When a class is immutable in the first place, its developers are forced
+to keep it cohesive and small. Because they can't encapsulate
+too much and they can't modify what's encapsulated. Just two or three
+arguments of a constructor and the reasonable limit is reached. Everything
+on top of that will look strange and clumsy.
+
+The immutable design of \code{jcabi-email} implements exactly the same
+email sending functionality, but employs seven classes for that, instead
+of one. Of course, the cohesiveness of each of them is much higher than
+the one of \code{SimpleEmail}. The length of each of them is below 300
+lines of code, which by itself is a perfect indicator of high readability
+and maintainability.
+Moreover, to extend the functionality of the library, existing classes
+don't need to be modified. Each new feature must be added through
+creation of new classes and implementing existing interfaces.
+
+Emprical results collected above confirm the logic explained. Larger classes
+tend to be mutable, which makes it easier for their authors to make them
+even larger when it's necessary.
 
